---
title: "hw02"
author: "Sam Shepard"
date: "2/4/2021"
output:
  github_document: default
  html_document: default
---

<!-- BMW: Please knit twice. Once to github_document and once to html_document. --> 

## Directions etc.

Assignment 02: Explore gapminder and use dplyr

<!-- BMW: In markdown, if you want things on new lines, you need to put blank lines between them. A single linebreak is ignored. -->

Jan 1, 0001

Objective

The goal is to practice with ggplot2 and dplyr to explore a dataset.


Setup

Work with the gapminder::gapminder data we explored in class.

Load up the packages you will need for this assignment using library():

Tidy Submission

Please be sure to follow the instructions for making a tidy submission, found on the assignments home page.

Exercise 1: Basic dplyr

Use dplyr functions to achieve the following.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(gapminder)
library(palmerpenguins)
```

## 1.1

Use filter() to subset the gapminder data to three countries of your choice in the 1970’s.

```{r}
samcountries <- gapminder %>%
  filter(country %in% c("Cuba", "Vietnam", "China") & year > 1969 & year < 1981)
```


1.2

Use the pipe operator %>% to select “country” and “gdpPercap” from your filtered dataset in 1.1.

```{r}
samcountries %>%
  select(country, gdpPercap)
```

1.3

Make a new variable in gapminder for the change in life expectancy from the previous measurement. Filter this table to show all of the entries that have experienced a drop in life expectancy.

Hint: you might find the lag() or diff() functions useful.

```{r}
gapminder %>% 
  mutate(changeLifeExp = lifeExp - lag(lifeExp)) %>% 
  filter(changeLifeExp < 0)
```

<!-- BMW: Almost. You computed a change in lifeExp, but didn't filter to show just the rows where lifeExp dropped. See above for a clean approach to the full project. -->

1.4

Filter gapminder so that it shows the max GDP per capita experienced by each country.

Hint: you might find the max() function useful here.

```{r}
gapminder %>% 
  group_by(country) %>% # BMW: For readability, always start a new line after %>% 
  filter(max == max (gdpPercap)) # BMW: The question asked you to _filter_ the dataset, not just show the summaries
```

1.5

Produce a scatterplot of Canada’s life expectancy vs. GDP per capita using ggplot2, without defining a new variable. That is, after filtering the gapminder data set, pipe it directly into the ggplot() function. In your plot, put GDP per capita on a log scale.

```{r}
gapminder %>%
  filter(country == "Canada") %>%
  ggplot()+
  aes(x = gdpPercap, y = lifeExp) +
  geom_point() +
  scale_x_log10(labels = scales::comma_format())
```

Exercise 2: Explore two variables with dplyr and ggplot2

Use palmerpenguins::penguins or another dataset of your choice. (Check out a dataset from the datasets R package if you want!)

2.1

Pick two quantitative variables to explore.

Make a summary table of descriptive statistics for these variables using summarize(). 
Include whatever staistics you feel appropriate (mean, median sd, range, etc.).

```{r}
penguins %>%
summarize(mean_bill_length = mean(bill_length_mm, na.rm = TRUE), # BMW: You probably want na.rm = TRUE
          mean_bill_depth = mean(bill_depth_mm, na.rm = TRUE),
          sd_bill_length = sd(bill_length_mm, na.rm = TRUE),
          sd_bill_depth = sd(bill_depth_mm, na.rm = TRUE)
          )
```

Make a scatterplot of these variables using ggplot().

```{r}
ggplot(penguins) +
  aes(x = bill_length_mm, y = bill_depth_mm, color = species) +
  geom_point(alpha = 0.2)
```

2.2

Pick one categorical variable and one quantitative variable to explore.

Make a summary table giving the sample size (hint: n()) and descriptive statistics for the quantitative variable by group.

```{r}
penguins %>%
  group_by(sex) %>%
  summarise(mean = mean(body_mass_g), # BMW: What were you were trying to do with tibble() there?
            n = n()
  )
```

Make one or more useful plots to visualize these variables.

```{r}
ggplot(penguins) +
  aes(sex, body_mass_g) + 
  geom_jitter(aes(color = sex), height = 0, width = .2) + # BMW: Always better to show both the raw data (points) and the summary (bar/boxplot)
  geom_boxplot(alpha = 0.2)
```
