---
title: "hw02"
author: "Sam Shepard"
date: "2/4/2021"
output:
  github_document: default
  html_document: default
---

## Directions etc.

Assignment 02: Explore gapminder and use dplyr
Jan 1, 0001
Objective
The goal is to practice with ggplot2 and dplyr to explore a dataset.

Setup
Work with the gapminder::gapminder data we explored in class.

Load up the packages you will need for this assignment using library():

Tidy Submission
Please be sure to follow the instructions for making a tidy submission, found on the assignments home page.

Exercise 1: Basic dplyr
Use dplyr functions to achieve the following.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```

## 1.1
Use filter() to subset the gapminder data to three countries of your choice in the 1970’s.

```{r}
#I can get 1 country and the year to filter, but not the countries together...
gapminder %>%
  filter(country == "Cuba" & country == "Vietnam" & country == "China", year > 1969 & year < 1981)

#also tried this:
gapminder %>%
  filter(country == "Cuba", "Vietnam", "China", year > 1969 & year < 1981)
```


1.2
Use the pipe operator %>% to select “country” and “gdpPercap” from your filtered dataset in 1.1.

```{r}
gapminder %>%
  select(country,gdpPercap)
```

1.3
Make a new variable in gapminder for the change in life expectancy from the previous measurement. Filter this table to show all of the entries that have experienced a drop in life expectancy.

Hint: you might find the lag() or diff() functions useful.

```{r}
library(Hmisc)

gapminder$lifeExp_lag <- Lag(gapminder$lifeExp)

  gapminder %>% 
  mutate(lifeExp_lag_final = lifeExp - lifeExp_lag, .keep = "all")
```

1.4
Filter gapminder so that it shows the max GDP per capita experienced by each country.

Hint: you might find the max() function useful here.

```{r}
gapminder %>% group_by(country) %>%
  summarise(max = max (gdpPercap))
```

1.5
Produce a scatterplot of Canada’s life expectancy vs. GDP per capita using ggplot2, without defining a new variable. That is, after filtering the gapminder data set, pipe it directly into the ggplot() function. In your plot, put GDP per capita on a log scale.

```{r}
gapminder %>%
  filter(country == "Canada") %>%
  ggplot()+
  aes(gdpPercap, lifeExp)+
  geom_point()+
  scale_x_log10(labels = scales::comma_format())
```