---
title: "hw03"
author: "Sam Shepard"
date: "1/25/2021"
output:
  github_document: default
---

# Homework 03

You will analyze data looking at the relationship between **green reputation** and three personality traits--**compassion**, **intellectual curiosity**, and **openness to experiences**.
The dataset includes data from **students** and **non-students**.

```{r}
library(dplyr)
library(tidyverse)
```

```{r}
dictionary <- readr::read_csv("green_dictionary.csv")
green_data <- readr::read_csv("green_data.csv")
```

For your assignment, do the following.

  1. Inspect the item responses (e.g., with graphs or by summarizing distinct values). 
  
```{r}
summary(green_data)
```
  
     Is anything unusual?
Missing values are coded as -99; below is the code to change the -99 values to "missing."
```{r}
green_data %>% 
mutate(
  across(.cols = everything (),
         ~ na_if(.x, "-99")
  ))

#Notes from class:
#everything = ALL COLUMNS
#~ = opens up for using ".x"
#.x = the current column
```

  2. Compute total scores for the four scales.

```{r}
green_data %>% 
  rowwise() %>% 
  mutate(
          GreenRep = mean(c(green1, green2r, green3, green4, green5), na.rm = TRUE),
          Compassion = mean(c(comp1, comp2, comp3, comp4, comp5, comp6r, comp7r, comp8r, comp9r, comp10r), na.rm = TRUE))
```



          IntCur = mean(c_across(intel1:intel10), na.rm = TRUE),
          Open = mean(c_across(open1:open10), na.rm = TRUE)
         )
```


##Recode variables as needed.
     
     
#Reverse-coding and creating total scores   
```{r}
green_data %>% 
  mutate(
   green2r = recode(green2, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
   comp6r = recode(comp6, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
   comp7r = recode(comp7, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
   comp8r = recode(comp8, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
   comp9r = recode(comp9, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
   comp10r = recode(comp10, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
   intel7r = recode(intel7, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
   intel8r = recode(intel8, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
   intel9r = recode(intel9, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
   intel10r = recode(intel10, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
   open7r = recode(open7, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
   open8r = recode(open8, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
   open9r = recode(open9, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"),
   open10r = recode(open10, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1")) %>% 
   
  rowwise() %>% 
  mutate(
          GreenRep = mean(green1, green2r, green3, green4, green5, trim = 0, na.rm = TRUE),
          Compassion = mean(comp1, comp2, comp3, comp4, comp5, comp6r, comp7r, comp8r, comp9r, comp10r, trim = 0, na.rm = TRUE),
          IntCur = mean(intel1, intel2, intel3, intel4, intel5, intel6, intel7r, intel8r, intel9r, intel10r, trim = 0, na.rm = TRUE),
          Open = mean(open1, open2, open3, open4, open5, open6, open7r, open8r, open9r, open10r, trim = 0, na.rm = TRUE)) %>% 
        
  select(id, green2, green2r, comp6, comp6r, intel7, intel7r, open7, open7r, GreenRep, Compassion, IntCur, Open)
  
#The table that populates doesn't show the decimals for the total scores?

```



```{r}
#recode categorical variable
green_data %>%
  mutate(
    student = recode(student, "1" = "not a student", "2" = "student")
  ) %>%
  select(id, student)
```


  3. Rescale the variables so that they go from 0-100 instead of the original range.
  
```{r}



```
    
    Name the rescaled variables `*_pomp`.
     
```{r}



```
  4. Make plots that illustrate the distributions of the 4 POMP-scored variables.
  
```{r}



```
  5. Make scatterplots showing the relationships between **green reputation** and each personality trait.
  
```{r}



```
     Include trend lines for **students** and **non-students**.

```{r}



```
     What do these plots show?
     
     
  6. Compare **green reputation** for students and non-students using a **rainfall plot** (bar + density + data points).
  
```{r}



```
 
  7. Compute a summary table of means, SDs, medians, minima, and maxima for the four total scores for students and non-students.
  
```{r}
green_data %>% 
  group_by(student) %>% 
  summarize(GreenRep_M = mean(GreenRep),
            GreenRep_SD = sd(GreenRep),
            GreenRep_Med = median(GreenRep),
            GreenRep_Min = min(GreenRep),
            GreenRep_Max = max(GreenRep),
            Compassion_M = mean(Compassion),
            Compassion_SD = sd(Compassion),
            Compassion_Med = median(Compassion),
            Compassion_Min = min(Compassion),
            Compassion_Max = max(Compassion),
            IntCur_M = mean(IntCur),
            IntCur_SD = sd(IntCur),
            IntCur_Med = median(IntCur),
            IntCur_Min = min(IntCur),
            IntCur_Max = max(IntCur),
            Open_M = mean(Open),
            Open_SD = sd(Open),
            Open_Med = median(Open),
            Open_Min = min(Open),
            Open_Max = max(Open)) %>% 
  ungroup()
```
  
In your assignment, prepare an RMarkdown file that includes both the coding and (for 4-7) narrative descriptions of your results.
