---
title: "hw03"
author: "Sam Shepard"
date: "1/25/2021"
output:
  github_document: default
---

# Homework 03

You will analyze data looking at the relationship between **green reputation** and three personality traits--**compassion**, **intellectual curiosity**, and **openness to experiences**.
The dataset includes data from **students** and **non-students**.

```{r}
library(dplyr)
library(tidyverse)
```

```{r}
dictionary <- readr::read_csv("green_dictionary.csv")
green_data <- readr::read_csv("green_data.csv")
```

For your assignment, do the following.

  1. Inspect the item responses (e.g., with graphs or by summarizing distinct values). 
  
```{r}
summary(green_data)
```
  
     Is anything unusual?
Missing values are coded as -99; below is the code to change the -99 values to "missing."
```{r}
green_data %>% 
mutate(
  across(.cols = everything (),
         ~ na_if(.x, "-99")
  ))

#Notes from class:
#everything = ALL COLUMNS
#~ = opens up for using ".x"
#.x = the current column
```

  2. Compute total scores for the four scales.

```{r}

green_data %>% 
  rowwise() %>% 
  mutate(GreenRep = mean(c(green1, green2, green3, green4, green5), na.rm = TRUE),
         )
```


     Recode variables as needed.
     
```{r}
#reverse-code variables
?across
green2
comp6:comp10
intel7:intel10
open7:open10

green_data %>% 
  mutate(
     across("green2", "comp6", "comp7", "comp8", "comp9", "comp10", "intel7", "intel8, "intel9", "intel10", "open7", "open8", "open9", "open10"),
    recode(.x, "1" = 5, "2" = 4, "3" = 3, "4" = 2, "5" = 1),
      ))
      
?across


green_data %>% 
  mutate(
   green2 = recode(green2, "1" = "5", "2" = "4", "3" = "3", "4" = "2", "5" = "1"))
```



```{r}
#recode categorical variable
green_data %>%
  mutate(
    student = recode(student, "1" = "not a student", "2" = "student")
  ) %>%
  select(id, student)
```


  3. Rescale the variables so that they go from 0-100 instead of the original range.
  
```{r}



```
    
    Name the recaled variables `*_pomp`.
     
```{r}



```
  4. Make plots that illustrate the distributions of the 4 POMP-scored variables.
  
```{r}



```
  5. Make scatterplots showing the relationships between **green reputation** and each personality trait.
  
```{r}



```
     Include trend lines for **students** and **non-students**.

```{r}



```
     What do these plots show?
     
     
  6. Compare **green reputation** for students and non-students using a **rainfall plot** (bar + density + data points).
  
```{r}



```
 
  7. Compute a summary table of means, SDs, medians, minima, and maxima for the four total scores for students and non-students.
  
```{r}
green_data %>% 
  group_by(student) %>% 
  summarize(GreenRep_M = mean(GreenRep),
            GreenRep_SD = sd(GreenRep),
            GreenRep_Med = median(GreenRep),
            GreenRep_Min = min(GreenRep),
            GreenRep_Max = max(GreenRep),
            Compassion_M = mean(Compassion),
            Compassion_SD = sd(Compassion),
            Compassion_Med = median(Compassion),
            Compassion_Min = min(Compassion),
            Compassion_Max = max(Compassion),
            IntCur_M = mean(IntCur),
            IntCur_SD = sd(IntCur),
            IntCur_Med = median(IntCur),
            IntCur_Min = min(IntCur),
            IntCur_Max = max(IntCur),
            Open_M = mean(Open),
            Open_SD = sd(Open),
            Open_Med = median(Open),
            Open_Min = min(Open),
            Open_Max = max(Open)) %>% 
  ungroup()
```
  
In your assignment, prepare an RMarkdown file that includes both the coding and (for 4-7) narrative descriptions of your results.
