---
title: "hw03"
author: "Sam Shepard"
date: "1/25/2021"
output:
  html_document:
    df_print: paged
  github_document: default
---

# Homework 03

### You will analyze data looking at the relationship between **green reputation** and three personality traits--**compassion**, **intellectual curiosity**, and **openness to experiences**.
## The dataset includes data from **students** and **non-students**.

```{r}
library(dplyr)
library(tidyverse)
library(graphics)
library(ggplot2)
```

```{r}
dictionary <- readr::read_csv("green_dictionary.csv")
green_data <- readr::read_csv("green_data.csv")
```

### 1. Inspect the item responses (e.g., with graphs or by summarizing distinct values). 
  
```{r}
summary(green_data)
```
  
### Is anything unusual?
**Missing values are coded as -99; below is the code to change the -99 values to "missing."**

```{r}
green_data_missing <- green_data %>% 
mutate(
  across(.cols = everything (),
         ~ na_if(.x, "-99")
  ))

#Notes from class:
#everything = ALL COLUMNS
#~ = opens up for using ".x"
#.x = the current column
```

### 2. Reverse-coding, recoding categorical variable, and creating total scores.
```{r}
reversed <- dictionary %>% 
  filter(Keying == -1|Keying == -2) %>% 
  pull(Item)
```

```{r}

green_data_missing_r <- green_data_missing %>% 
  mutate(
    across(all_of(reversed),
           ~ 6 - .x,
           .names = "{.col}r"),
     student = recode(student, "1" = "not a student", "2" = "student", .missing = "(Unknown)")
  )
```




```{r}
green_data_missing_r_totals <- green_data_missing_r %>% 
  rowwise() %>% 
  mutate(
          GreenRep = mean(c_across(c(green1, green2r, green3:green5)), trim = 0, na.rm = TRUE),
          Compassion = mean(c_across(c(comp1:comp5, comp6r:comp10r)), trim = 0, na.rm = TRUE),
          IntCur = mean(c_across(c(intel1:intel6, intel7r:intel10r)), trim = 0, na.rm = TRUE),
          Open = mean(c_across(c(open1:open6, open7r:open10r)), trim = 0, na.rm = TRUE)) %>% 
        
  select(id, student, GreenRep, Compassion, IntCur, Open)
```


### 3. Rescale the variables so that they go from 0-100 instead of the original range. Name the rescaled variables `*_pomp`.

```{r}
(green_data_scored_POMP <- green_data_missing_r_totals %>% 
mutate(
GreenRep_pomp = (GreenRep * 20),
Compassion_pomp = (Compassion * 20),
IntCur_pomp = (IntCur * 20),
Open_pomp = (Open * 20)
) %>% 

select(id, student, GreenRep_pomp, Compassion_pomp, IntCur_pomp, Open_pomp))

```
    
### 4. Make plots that illustrate the distributions of the 4 POMP-scored variables.

**Overall, participants are high in compassion, intellectual curiosity, and openness to experience. The data for green reputation is more normally distributed.**
  
```{r}
ggplot(green_data_scored_POMP, aes(x=GreenRep_pomp)) +
            geom_histogram()

ggplot(green_data_scored_POMP, aes(x=Compassion_pomp)) +
            geom_histogram()

ggplot(green_data_scored_POMP, aes(x=IntCur_pomp)) +
            geom_histogram()

ggplot(green_data_scored_POMP, aes(x=Open_pomp)) +
            geom_histogram()
```
  
  

### 5. Make scatterplots showing the relationships between **green reputation** and each personality trait. Include trend lines for **students** and **non-students**.

**Write narrative here once the student lines are figured out**

**Compassion & Green Reputation**  
```{r}
green_data_scored_POMP %>% 
ggplot() +
  aes(GreenRep_pomp, Compassion_pomp, linetype = student) +
  geom_point() +
  geom_jitter() +
  geom_smooth()

#I tried geom_line(), but that turned out absolutely horrific, so I went with geom_smooth for this draft.
```

**Intellectual Curiosity & Green Reputation**  

```{r}
green_data_scored_POMP %>% 
ggplot() +
  aes(GreenRep_pomp, IntCur_pomp, linetype = student) +
  geom_point() +
  geom_jitter() +
  geom_smooth()

```

**Openness to Experience & Green Reputation** 

```{r}
green_data_scored_POMP %>% 
ggplot() +
  aes(GreenRep_pomp, Open_pomp, linetype = student, color = student) +
  geom_point() +
  geom_jitter() +
  geom_smooth()

```





     
### 6. Compare **green reputation** for students and non-students using a **rainfall plot** (bar + density + data points).

**The rainfall plot indicates that the distributions for students and non-students are very similar.**
  
```{r}
green_data_scored_POMP %>%
ggplot() +
  aes(y = student, x = GreenRep_pomp, na.omit, color = student) +
  geom_jitter() +
  geom_boxplot(alpha = .1) +
  ggdist::stat_slab()
  
```
 
### 7. Compute a summary table of means, SDs, medians, minima, and maxima for the four total scores for students and non-students.

**Write narrative description after figuring out what is wrong with this summary table**
  
```{r}
green_data_scored_POMP %>% 
  group_by(student) %>% 
  summarize(GreenRep_M = mean(GreenRep_pomp),
            GreenRep_SD = sd(GreenRep_pomp),
            GreenRep_Med = median(GreenRep_pomp),
            GreenRep_Min = min(GreenRep_pomp),
            GreenRep_Max = max(GreenRep_pomp),
            Compassion_M = mean(Compassion_pomp),
            Compassion_SD = sd(Compassion_pomp),
            Compassion_Med = median(Compassion_pomp),
            Compassion_Min = min(Compassion_pomp),
            Compassion_Max = max(Compassion_pomp),
            IntCur_M = mean(IntCur_pomp),
            IntCur_SD = sd(IntCur_pomp),
            IntCur_Med = median(IntCur_pomp),
            IntCur_Min = min(IntCur_pomp),
            IntCur_Max = max(IntCur_pomp),
            Open_M = mean(Open_pomp),
            Open_SD = sd(Open_pomp),
            Open_Med = median(Open_pomp),
            Open_Min = min(Open_pomp),
            Open_Max = max(Open_pomp))

#Most of the values in this summary table read, "NA" or "NaN"
```
