---
title: "Class Notes 2-11-21"
author: "Sam Shepard"
date: "2/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reviewing/trouble shooting/questions
#Select = select columns
#filter = select rows
```{r}
library(tidyverse)
library(palmerpenguins)

#Putting variables on x vs. y variables - use aes and just specify by typing out "x = " and "y = "
ggplot(penguins) +
  aes(x = species, y = flipper_length_mm)+
  geom_point()+
  theme_classic()
#or put "geom_jitter" to spread out points
```

#Renaming x and y axes:
```{r}
#use scale functions
#discrete = change categorical variable
ggplot(penguins) +
  aes(x = species, y = flipper_length_mm)+
  geom_point()+
  scale_x_discrete("Species")+
  theme_classic()
```

#Renaming axes before plotting
```{r}
penguins %>% 
  mutate(sex = recode(sex, female = "Female", male = "Male"))
ggplot(penguins) +
  aes(x = species, y = flipper_length_mm,
      color = sex)+ 
  geom_point()+
  scale_x_discrete("Species")+
  theme_classic()
```
##Changing axis title
#can ALSO use xlab, ylab, or labs(x = "Species")
#scale functions just give you more things to work with so it's a good intro/function to start with


##Colors!
```{r}
penguins %>% 
  mutate(sex = recode(sex, female = "Female", male = "Male"))
ggplot(penguins) +
  aes(x = species, y = flipper_length_mm,
      color = sex)+
  geom_point()+
  scale_x_discrete("Species")+
  scale_color_brewer(palette = "Blues")+
  theme_classic()
```

##How to get good at programming
#Think through it stepwise
#first thing you think about - what are the specific things I want to do?
#Break up steps into sub-steps
#think through and write out each step!



###Dplyr
#https://talks.andrewheiss.com/2020-gpl/01_welcome-tidyverse/04_transform-data.html
#Main verbs:
#filter
#First argument: data frame to transform
#"...": which rows of data do you want to match

```{r}
filter(.data = gapminder, country == "Denmark")
```

#Important info about choosing more than 1 type of "row"
```{r}
#BAD: "condition ==c("A", "B")
#GOOD: "condition%in%c("A", "B")
#above - saying test all values against each row

#==, >=, <= all compare 1 element to 1 element

#%in% - compare 1 element to multiple different elements

#%in% is a relational operator, just like the other "normal" ones

#"=" is not an operator because it assigns values
```

##Commmon Mistakes
#Using = rather than ==
#Quote use

Extract rows that pass every test:

```{r}
filter(gapminder, country == "Denmark", year > 2000)
#comma - same as &
#how to use the "OR" operation - button above the "Enter" button
#"!=" - excludes
```

#Common syntax
verb(data, ...)

verb = dplyr function
data = data frame
... = stuff you want to do


mutate(.data, columns you want to make)
```{r}
mutate(gapminder, gdp = gdpPercap * pop)
#new column - gdp
#multiply the two existing columns to create new vbl gdp
```


```{r}
#create 2 vbls at same time
#this isn't working for some reason?
mutate(gapminder, gdp = gdpPercap * pop,
       pop_mil = round(pop / 1000000)
```
#Do things for some rows and other things for other rows
ifelse(test,
       value if true
       value if false)
```{r}
#Label data for before or after 1960
mutate(gapminder,
       after_1960 = ifelse (year>1960, TRUE, FALSE))
```

#What if you have multiple verbs?
#piping!!!

#Summarize
```{r}
gapminder %>%  summarize(mean_life = mean(lifeExp))
#has same structure has mutate
# fyi - data frames do not change unless you assign them to another name
```

#Group_by
```{r}
#summarize becomes most useful when combined with group_by
gapminder %>% group_by(continent) %>% #nothing happens when you run only this command
summarize(n_countries = n_distinct("country"))
```
