---
title: "2/9/21"
author: "Sam Shepard"
date: "2/9/2021"
output: html_document
---

## Class Notes 2/9/21
If code isn't working, make sure you don't have competing packages loaded.


Long version:
```{r}
gapminder %>%
  summarize(mean_pop = mean(pop),
          (sd_pop = sd(pop),
           (mean_gdp = mean(gdpPercacp),
           sd_gdp = sd(gdpPercap)
           )
```
            
            
```{r}            
gapminder%>%
  summarize(across(c(pop, gdpPercap),
                   list (Mean = mean, SD = sd)))
```
  
DRY - don't repeat yourself

```{r}

```
  
"~" lets you add arguments that are outside the defualts; means that you're going to write out the whole function, rather than just the call name 
".x" can be the placeholder of the variable you're working with
```{r}
gapminder %>% 
  group_by(continent) %>% 
  summarize(across(c(pop, gdpPercap),
                   list(Mean = ~mean(.x, na.rm = TRUE),
                        SD = ~sd(.x, na.rm = TRUE))))
```  
  
  
Rowwise command
```{r}
psych::bfi %>% 
  rowwise() %>% 
  mutate(A_tot = mean(c(A1, A2, A3, A4, A5), na.rm = TRUE),
         )
``` 
mutate - creates a new column 
c = creates vector

1. Use the `psych::bfi` dataset.
  :: pulls out bfi data from psych package
  package::object
  
2. Compute mean scores for each of the Big Five traits for each person.
3. Keep just Big Five scores, gender, age, and education.
4. Summarize the mean, standard deviation, minimum, and maximum values for the Big Five scores separately by gender groups.
5. Summarize the mean, standard deviation, minimum, and maximum values for the Big Five scores separately by educational level groups.
```{r}
bfi_scored <- psych::bfi %>%
rowwise() %>% 
mutate(Agree = mean(c(A1,A2,A3,A4,A5), na.rm = TRUE),
       Con = mean(c_across(C1:C5), na.rm = TRUE),
       Extra = mean(c_across(E1:E5), na.rm = TRUE),
       Neuro = mean(c_across(N1:N5), na.rm = TRUE),
       Open = mean(c_across(O1:O5), na.rm = TRUE),
       ) %>% 
  select(gender,education, age, Agree, Con, Open, Extra, Neuro) %>% 
  

bfi_scored %>% 
  group_by(gender) %>% 
  summarize(across(Agree:Neuro,
            list(M = mean, SD = sd, min = min, max = max)
            )
)
```  
c_across - makes a vector
functions need to be inside "across" when you use "across"
don't need "c_across" when you use ":"
help function:
the first thing that needs to be after across is the columns
2nd - list of functions you want to use

R does not care about new lines! You can put paratheses into new lines to help you make sure you have the correct number of parentheses

  
```{r}

``` 
  
  
            