---
title: "Beer Advocate: ABV and quality ratings"
author: "Your Name Here"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

For this activity, download the data file at this location to the `data` subfolder in your `participation` folder. 
This dataset contains information about different catagories of beer as rated on <beeradvocate.com>.
The data are originally from [here](http://scottjanish.com/relationship-of-abv-to-beer-scores/).


read.csv is base R and worse to use; read_csv tells you data type etc. so it better to use.
COLORS ARE IMPORTANT
backticks refer to columns in which there are a space character in them
file - edit - code - clean up code
ctrl i = clean up code
```{r}
url <- "https://raw.githubusercontent.com/USF-Psych-DataSci/Classroom/master/data/beer-advocate.csv"

beer_advocate <- read_csv(url, na = c("Non-alcoholic", 'NA', ""),
  col_types = cols(
  `Beer Category` = col_character(),
  `Average ABV` = col_double(),
  rAVG = col_double()
))


```

The dataset contains 3 variables:

  1. Beer category
  2. Average alcohol by volume (ABV) in percentage (%)
  3. Average quality rating (rAVG) on a 1–5 scale
  
You should write a script to do the following:

  1. Download the dataset with an appropriate filename.
  2. Import the dataset into R.
     Be sure to check for any importing problems and correct as needed.
  3. Make a figure illustrating the relationship between ABV and beer quality.
  
```{r}
ggplot(data = beer_advocate,
       aes(x = `Average ABV`, y = rAVG))+
      geom_point()
      
```

  4. Compute a summary table with the mean and SD of each variable, along with their correlation.
  
```{r}
summarise(beer_advocate,
          across(c(`Average ABV`, rAVG),
                 list(mean = ~mean(.x, na.rm = TRUE),
                      sd = ~sd(.x, na.rm = TRUE))
                 
          ),
                cor = cor(`Average ABV`, rAVG, use = "na.or.complete")
                 
                 
                 
          )
          


```


  5. Describe the results.
